---
title: "In-class project"
subtitle: "EC 421"
author: "Edward Rubin"
date: "Winter 2021"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    # self_contained: true
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, middle

```{r, setup, include = F}
# devtools::install_github("dill/emoGG")
library(pacman)
p_load(
  broom, here, tidyverse,
  emoGG, latex2exp, ggplot2, ggthemes, ggforce, viridis, extrafont, gridExtra,
  wooldridge,
  kableExtra, snakecase,
  data.table, dplyr, estimatr,
  lubridate,
  magrittr, knitr, parallel
)
# Define pink color
red_pink <- "#e64173"
turquoise <- "#20B2AA"
orange <- "#FFA500"
red <- "#fb6107"
blue <- "#2b59c3"
green <- "#8bb174"
grey_light <- "grey70"
grey_mid <- "grey50"
grey_dark <- "grey20"
purple <- "#6A5ACD"
# Dark slate grey: #314f4f
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  warning = F,
  message = F
)
opts_chunk$set(dev = "svg")
options(device = function(file, width, height) {
  svg(tempfile(), width = width, height = height)
})
options(crayon.enabled = F)
options(knitr.table.format = "html")
# Column names for regression results
reg_columns <- c("Term", "Est.", "S.E.", "t stat.", "p-Value")
# Function for formatting p values
format_pvi <- function(pv) {
  return(ifelse(
    pv < 0.0001,
    "<0.0001",
    round(pv, 4) %>% format(scientific = F)
  ))
}
format_pv <- function(pvs) lapply(X = pvs, FUN = format_pvi) %>% unlist()
# Tidy regression results table
tidy_table <- function(x, terms, highlight_row = 1, highlight_color = "black", highlight_bold = T, digits = c(NA, 3, 3, 2, 5)) {
  x %>%
    tidy() %>%
    select(1:5) %>%
    mutate(
      term = terms,
      p.value = p.value %>% format_pv()
    ) %>%
    kable(
      col.names = reg_columns,
      escape = F,
      digits = digits
    ) %>%
    kable_styling(font_size = 20) %>%
    row_spec(1:nrow(tidy(x)), background = "white") %>%
    row_spec(highlight_row, bold = highlight_bold, color = highlight_color)
}
```

# Prologue

---
name: schedule

# Schedule

## Last Time

Instrumental variables

## Today

Extra-credit prediction competition
<br>.note[Submissions due by midnight (Pacific) tonight.]

## Upcoming

- **Last problem** set due by midnight (PST) on Friday 
- **Review** in lecture on Thursday (review materials are on Canvas)
- **Final** next week: Canvas .hi[8AM Pacific, Tuesday, 16 March 2021]

---
layout: false
class: inverse, middle
# Prediction

---
# Prediction
## Intro

Most tasks in econometrics boil down to one of two goals:

$$
\begin{align}
  y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \cdots + \beta_k x_k + u
\end{align}
$$

--

1. .hi-purple[Prediction:] Accurately and dependably .purple[predict/forecast] $\color{#6A5ACD}{y}$ using on some set of explanatory variables—doesn't need to be $x_1$ through $x_k$. Focuses on $\color{#6A5ACD}{\hat{y}}$. $\beta_j$ doesn't really matter.

--

1. .hi[Causal estimation:].super[.pink[†]] Estimate the actual data-generating process—learning about the true, population model that explains .pink[how] $\color{#e64173}{y}$ .pink[changes when we change] $\color{#e64173}{x_j}$—focuses on $\color{#e64173}{\beta_j}$. Accuracy of $\hat{y}$ is not important.

.footnote[
.pink[†] Often called *causal identification*.
]

---
# Prediction
## Competition

Today your job is to figure out the model that best .hi[predicts] the outcome `y`.

**Specifics**

- **Train** (build) your model using the `train.csv` dataset.
  - 15 predictors (regressors): `x1`, `x2`, ... `x15`
  - outcome: `y`
- **Predict** the outcome for the `test.csv` data set. 
- **Submit** a CSV of your predictions. <br> The CSV should only have a column of your preditions.

**Reward:** Better preditions = more extra-credit points.

---
# Prediction
## In R

Fit a model and then use `predict()`  to predict onto `newdata`.

```{r, predict, eval = F}
# Load packages
library(pacman)
p_load(tidyverse)
# Load datasets
train_df = read_csv("train.csv")
test_df = read_csv("test.csv")
# Fit a model
my_model = lm(y ~ x1 + x15 + x1:x15, data = train_df)
# Predict onto new dataset
my_predictions = predict(my_model, newdata = test_df)
# Save as CSV
write_csv(
  as.data.frame(my_predictions),
  "my-predictions.csv"
)
```

---
exclude: true

```{r, print pdfs, echo = F, eval = F}
pagedown::chrome_print(
  input = "prompt-02.html",
  timeout = 60
)
```