---
title: "Data Science for Economists"
subtitle: "Randomized Controlled Trials and Simulations"
author: Kyle Coombs
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts,css/my-css.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
---

```{css, echo=FALSE}
@media print {
  .has-continuation {
    display: block !important;
  }
}
pre {
  max-height: 350px;
  overflow-y: auto;
}

pre[class] {
  max-height: 100px;
}
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE) 
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache=TRUE, fig.width = 8, fig.height = 6, fig.align='center')
if (!require("pacman")) install.packages("pacman")
devtools::install_github("dill/emoGG")
pacman::p_load(
  tidyverse, gganimate, estimatr, magick, directlabels, ggthemes, fixest, jtools,
  rdrobust, rddensity, scales, Cairo, broom, here, emoGG, latex2exp, ggplot2,
  viridis, extrafont, gridExtra, kableExtra, data.table, dplyr, lubridate,
  magrittr, knitr, parallel, transformr
)

# Define pink color
red_pink <- "#e64173"
turquoise <- "#20B2AA"
grey_light <- "grey70"
grey_mid <- "grey50"
grey_dark <- "grey20"
purple <- "#6A5ACD"
# Dark slate grey: #314f4f
# Knitr options
# A blank theme for ggplot
theme_empty <- theme_bw() + theme(
  line = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  axis.text = element_blank(),
  plot.title = element_blank(),
  axis.title = element_blank(),
  plot.margin = structure(c(0, 0, -0.5, -1), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
theme_simple <- theme_bw() + theme(
  line = element_blank(),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  axis.text.x = element_text(size = 18, family = "STIXGeneral"),
  axis.text.y = element_blank(),
  axis.ticks = element_blank(),
  plot.title = element_blank(),
  axis.title = element_blank(),
  # plot.margin = structure(c(0, 0, -1, -1), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
theme_axes_math <- theme_void() + theme(
  text = element_text(family = "MathJax_Math"),
  axis.title = element_text(size = 22),
  axis.title.x = element_text(hjust = .95, margin = margin(0.15, 0, 0, 0, unit = "lines")),
  axis.title.y = element_text(vjust = .95, margin = margin(0, 0.15, 0, 0, unit = "lines")),
  axis.line = element_line(
    color = "grey70",
    size = 0.25,
    arrow = arrow(angle = 30, length = unit(0.15, "inches")
  )),
  plot.margin = structure(c(1, 0, 1, 0), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
theme_axes_serif <- theme_void() + theme(
  text = element_text(family = "MathJax_Main"),
  axis.title = element_text(size = 22),
  axis.title.x = element_text(hjust = .95, margin = margin(0.15, 0, 0, 0, unit = "lines")),
  axis.title.y = element_text(vjust = .95, margin = margin(0, 0.15, 0, 0, unit = "lines")),
  axis.line = element_line(
    color = "grey70",
    size = 0.25,
    arrow = arrow(angle = 30, length = unit(0.15, "inches")
  )),
  plot.margin = structure(c(1, 0, 1, 0), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
theme_axes <- theme_void() + theme(
  text = element_text(family = "Fira Sans Book"),
  axis.title = element_text(size = 18),
  axis.title.x = element_text(hjust = .95, margin = margin(0.15, 0, 0, 0, unit = "lines")),
  axis.title.y = element_text(vjust = .95, margin = margin(0, 0.15, 0, 0, unit = "lines")),
  axis.line = element_line(
    color = grey_light,
    size = 0.25,
    arrow = arrow(angle = 30, length = unit(0.15, "inches")
  )),
  plot.margin = structure(c(1, 0, 1, 0), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
theme_metro <- function(x) {
  theme_classic() + 
  theme(panel.background = element_rect(color = '#FAFAFA',fill='#FAFAFA'),
        plot.background = element_rect(color = '#FAFAFA',fill='#FAFAFA'),
        text = element_text(size = 16),
        axis.title.x = element_text(hjust = 1),
        axis.title.y = element_text(hjust = 1, angle = 0))
}
theme_void_metro <- function(x) {
  theme_void() + 
  theme(panel.background = element_rect(color = '#FAFAFA',fill='#FAFAFA'),
        plot.background = element_rect(color = '#FAFAFA',fill='#FAFAFA'),
        text = element_text(size = 16))
}
theme_metro_regtitle <- function(x) {
  theme_classic() + 
  theme(panel.background = element_rect(color = '#FAFAFA',fill='#FAFAFA'),
        plot.background = element_rect(color = '#FAFAFA',fill='#FAFAFA'),
        text = element_text(size = 16))
}

setFixest_dict(c('above_cutTRUE'='Above Cut','treatmentTRUE'='Treatment','treatedTRUE'='Treated','X_center:treatmentTRUE'='X_center x Treatment','treatedTRUE:I(X_centered^2)'='Treated x X_centered squared'))
setFixest_etable(digits=2,fitstat="n")
```

# Table of contents

- [Introduction to RCTs](#rcts)
- [Direct Rental Assistance Example](#rental-assistance)
- [Power Analysis](#power-analysis)
- [Simulation](#simulation)

---
name: rcts
class: inverse, center, middle

# Introduction to RCTs

---

# Introduction to RCTs

- Randomized Controlled Trials (RCTs) are a gold standard in research for establishing causal relationships.
- They allow researchers to control potential outcomes by randomly assigning subjects to treatment and control groups.
- This randomization helps ensure that any differences in outcomes can be attributed to the treatment itself.

---
name: rental-assistance
class: inverse, center, middle

# Direct Rental Assistance Example

---

# Direct Rental Assistance Example

- On problem set, several of note DRA could help people have more control over their rent and ability to move.

- That is a great idea and better yet it is empirically testable.

- Let's say we give every low-income household in the U.S. direct rental assistance.

- We can't just give it to all of them and look at what they do:
  
1. That's too expensive
2. We wouldn't see the counterfactual

- High income households are a bad counterfactual too

- Given we likely won't give it to everyone immediately (budget constraints):

- Why not randomize the stages?

---
class: white-slide

```{r, sample_size, echo = F}
xd <- 9
yd <- 6
sample_size <- xd * yd
emoji_type <- "1f4b5"
house_emoji <- "1f3e0"
periods <- 2
```

.hi-slate[`r sample_size` households of varying starting rents (the color of the squares)]
```{r, house_plot1, echo = F}
set.seed(123)
house_df <- expand.grid(x = 1:xd, y = 1:yd) %>%
  mutate(
    trt = sample(x =c(1,0), size = xd*yd, replace = T, prob = c(0.51, 0.49)),
    income = 50000 + 15000 * rnorm(xd*yd),
    age = round(45 + 12 * rnorm(xd*yd)),
    family_size = round(pmax(1, 2.5 + rnorm(xd*yd))),
    years_in_residence = round(pmax(0, 5 + 3 * rnorm(xd*yd)),1),
    black = sample(x = c(T, F), size = xd*yd, replace = T, prob = c(0.6, 0.4))
  ) %>%
  arrange(-y, x) %>%
  mutate(hh = 1:(xd*yd) %>% str_pad(2, "left", "0")) %>%
  arrange(y, x)

house_did <- rbind(
  house_df %>% mutate(t=1),
    house_df %>% mutate(t=2)
  ) %>%
    mutate(after=t-1,
      rent=3+x+y-trt-5*trt*after+2*after+rnorm(xd*yd))

house_plot <- ggplot(data = filter(house_did,after==0), aes(x, y,fill=rent)) +
  geom_tile(color = "grey40", size = 0.2) +
  geom_emoji(emoji = house_emoji, size=0.1) +
  scale_fill_viridis(option = "magma", direction = -1) +
  theme_void() +
  theme(legend.position = "none") +
  coord_equal()
house_plot
```

---
class: white-slide
count: false

.hi-orange[Randomly assign treatment] in period 1
```{r, house_plot2, echo = F}
house_plot +
  geom_emoji(data = filter(house_did,trt==1,after==0), emoji = emoji_type)
```

Give to the rest in period 2

---
# Balance test

- After a randomization, we can check if the treated and control groups are balanced.

- We can do this by looking at the distribution of the covariates between the two groups.

- If there is no balance, we may need to re-randomize.

- Easiest way is to do a two-sample t-test

```{r balance_test, echo = F}
house_df %>% select(rent,income,age,family_size,years_in_residence,black) %>% 
  map(~ t.test(.x ~ house_df$trt)) %>% 
  map_dfr(~ tidy(.), .id = "variable") %>% 
  mutate(estimate = -round(estimate,2),
    Difference = case_when(p.value<0.01 ~ paste0(estimate,"***"),
                           p.value<0.05 ~ paste0(estimate,"**"),
                           p.value<0.1 ~ paste0(estimate,"*"),
                           TRUE ~ as.character(estimate)),
                           Treated = estimate2,
                           Untreated = estimate1) %>% 
  select(variable, estimate, p.value, Treated, Untreated) %>%
  kable(caption = "Two-sample t-test balance test between treated and control households (hundreds of $)")
```



---
# Easiest analysis two-sample t-test

- Two-sample t-test is the easiest analysis to compare means between two groups

```{r two_sample_ttest, echo = T}
ttest <- t.test(rent ~ trt, data = filter(house_did,after==1))
ttest %>% 
  tidy() %>% 
  mutate(estimate = -round(estimate,2),
    Difference = case_when(p.value<0.01 ~ paste0(estimate,"***"),
                           p.value<0.05 ~ paste0(estimate,"**"),
                           p.value<0.1 ~ paste0(estimate,"*"),
                           TRUE ~ as.character(estimate)),
     Treated = estimate2,
     Untreated = estimate1) %>% 
  select(Difference, Untreated, Treated) %>% 
  kable(caption = "Two-sample t-test of neighborhood quality between treated and control households (hundreds of $)")
```

---
name: power-analysis
class: inverse, center, middle

# Power Analysis

---

# Power Analysis

- We have statistical significance, but how much can we trust it?

- How likely is it that this is a fluke and the true effect is actually zero? 
  - A false positive or Type I error, let's say $\alpha = Pr(\text{reject } H_0 | H_0 \text{ is true})$

- If we had failed to reject the null hypothesis of 0, how likely is it that the true effect is actually 0?
  - A false negative or Type II error, let's say $\beta = Pr(\text{fail to reject } H_0 | H_0 \text{ is false})$

- Statistical power is defined as: $1 - \beta$

- Power analysis helps understand the probability of these errors -- given some assumptions:

- We can use power analysis to determine the sample size required to detect an effect of a given size with a certain degree of confidence.

- Alternatively, we can use power analysis to determine the power of a given sample size to detect an effect of a given size.

---
# How does it work? 

- Scientists determine the type I error rate, $\alpha$ -- that's 

- But type II errors depend on a ton of factors:
  - Effect size
  - Sample size
  - Variability of the outcome
  - Type I error rate

---
# Effect: Magnitude of difference

```{r densities,echo=FALSE}
# Create two groups with different means
set.seed(123) # For reproducibility
tibble(
  id = 1:10000,
  trt = rep(c(0,1), each=5000),
  rent = rnorm(10000, mean = ifelse(trt==1, 12, 13), sd=2)
) -> small_treatment

# Create another dataset with more overlap
tibble(
  id = 1:10000, 
  trt = rep(c(0,1), each=5000),
  rent = rnorm(10000, mean = ifelse(trt==1, 8, 13), sd=2)
) -> big_treatment

bind_rows(
  small_treatment %>% mutate(plot = "Hard"),
  big_treatment %>% mutate(plot = "Easy")
) %>%
  mutate(trt = ifelse(trt==1, "Treatment", "Control")) %>%
  ggplot(aes(x=rent, color=trt)) +
  geom_density(alpha=0.5) +
  facet_wrap(~plot) +
  theme_metro() +
  labs(x="Rent", y="")
```

---
# Standard deviation: Variability of the outcome

```{r densities-sd,echo=FALSE}
# Create two groups with different means
set.seed(123) # For reproducibility
tibble(
  id = 1:10000, 
  trt = rep(c(0,1), each=5000),
  rent = rnorm(10000, mean = ifelse(trt==1, 8, 13), sd=5)
) -> high_sd_treatment

bind_rows(
  high_sd_treatment %>% mutate(plot = "High SD: Hard to tell difference"),
  big_treatment %>% mutate(plot = "Low SD: Easy to tell difference")
) %>%
  mutate(trt = ifelse(trt==1, "Treatment", "Control")) %>%
  ggplot(aes(x=rent, color=trt)) +
  geom_density(alpha=0.5) +
  facet_wrap(~plot) +
  theme_metro() +
  labs(x="Rent", y="")
```

---
# Sample Size:

```{r sample_size,echo=FALSE}
tibble(
  id = 1:10, 
  trt = rep(c(0,1), each=5),
  rent = rnorm(10, mean = ifelse(trt==1, 8, 13), sd=2)
) -> low_sample

bind_rows(
  low_sample %>% mutate(plot = "Low Sample (n=10)"),
  big_treatment %>% mutate(plot = "High Sample (n=10000)")
) %>%
  mutate(trt = ifelse(trt==1, "Treatment", "Control")) %>%
  ggplot(aes(x=rent, color=trt)) +
  geom_density(alpha=0.5) +
  facet_wrap(~plot) +
  theme_metro() +
  labs(x="Rent", y="")

```

---
# Type I error rate: $\alpha=0.05$ vs. $\alpha=0.001$

```{r type_i_error,echo=FALSE}
t_values <- seq(-4, 4, length.out = 1000)
t_dist <- dt(t_values, df = 30)  # df = degrees of freedom

# Define critical values for a two-tailed test at alpha = 0.05
critical_value <- qt(0.975, df = 30)
critical_value_0.001 <- qt(0.9995, df = 30)

# Create the plot
alpha0.05 <- ggplot(data.frame(t_values, t_dist), aes(x = t_values, y = t_dist)) +
  geom_area(fill = "grey", alpha = 0.5) +
  geom_area(data = subset(data.frame(t_values, t_dist), t_values < -critical_value),
            aes(x = t_values, y = t_dist), fill = "darkcyan", alpha = 0.8) +
  geom_area(data = subset(data.frame(t_values, t_dist), t_values > critical_value),
            aes(x = t_values, y = t_dist), fill = "darkcyan", alpha = 0.8) +
  geom_vline(xintercept = c(-critical_value, critical_value), linetype = "dashed", color = "grey") +
  labs(x = "t", y = NULL, title = "Type I error rate α = 0.05") +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

alpha0.001 <- ggplot(data.frame(t_values, t_dist), aes(x = t_values, y = t_dist)) +
  geom_area(fill = "grey", alpha = 0.5) +
  geom_area(data = subset(data.frame(t_values, t_dist), t_values < -critical_value_0.001),
            aes(x = t_values, y = t_dist), fill = "darkcyan", alpha = 0.8) +
  geom_area(data = subset(data.frame(t_values, t_dist), t_values > critical_value_0.001),
            aes(x = t_values, y = t_dist), fill = "darkcyan", alpha = 0.8) +
  geom_vline(xintercept = c(-critical_value_0.001, critical_value_0.001), linetype = "dashed", color = "grey") +
  labs(x = "t", y = NULL, title = "Type I error rate α = 0.001") +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

grid.arrange(alpha0.05, alpha0.001, nrow = 1)
```

---
# How do we calculate power?

The power calculations vary by test statistic, which can make them pretty tedious!

T-test:

$$
1-\beta \approx 1 - \Phi \left(1.64 - \frac{\text{Effect}}{\text{SD}/\sqrt{n}}\right)
$$

where $\Phi$ is the cumulative distribution function of the standard normal distribution.

You can invert that to solve for the sample size for a given power:

$$
n \approx \left(\frac{\text{Effect}}{\text{SD}}\right)^2 \left(1.64 + \Phi^{-1}(1-\beta)\right)^2
$$

---
# Using tools like `pwr`

Thankfully, you don't have to do this by hand! `pwr` is a package that can help you calculate power for a variety of tests (but not all of them).

```{r pwr, echo = T}
pwr::pwr.t.test(d = 0.5, power = 0.8, sig.level = 0.05, type = "two.sample", alternative = "two.sided")
```

- A software called G*Power can do a lot more, but it is limited by what is pre-programmed

- But what happens when it gets more complex?

---
# Power Analysis for more Complex Designs

- So before we did a two-way t-test and got a difference of

- But that may be insufficient if the households were not perfectly balanced too start.

- Plus, maybe household rents are correlated over time (need clustering)

- Then we'd do a diff-in-diff analysis

```{r diff_in_diff, echo = T}
feols(rent ~ after*trt | hh + t, data = house_did,cluster=~hh) %>% 
  etable() %>%
  kable(caption = "Diff-in-diff analysis of neighborhood\nquality between treated and control households (hundreds of $)")
```

---
# Power Analysis

- Uh oh, how do you do power analysis for diff-in-diff with clustering?

- The calculation is going to more complex

---







---


---
name: simulation-ai
class: inverse, center, middle

# Simulation

---

# Simulation Basics in R

- Simulations are a powerful tool for modeling complex systems and testing hypotheses.
- In R, you can generate random data using functions like `rnorm()`, `runif()`, and `rbinom()`.
- Simulations often involve running multiple iterations to assess variability and robustness.
- Results can be analyzed using summary statistics, visualizations, and statistical tests.

---

# AI in Simulation

- AI can enhance simulation by automating the writing and execution of simulation code.
- With carefully worded instructions, AI can generate complex simulations efficiently.
- This approach can improve the robustness and reliability of research findings by exploring a wider range of scenarios.

```{r gen_pdf, include = FALSE, cache = FALSE, eval = TRUE}
infile=knitr::current_input() %>% stringr::str_replace('.Rmd', '.html')
print(infile)
pagedown::chrome_print(input = infile, timeout = 100)
```
