---
title: "Data Science for Economists"
# subtitle: "<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>"
subtitle: "Lecture 5: Data cleaning & wrangling: (1) Tidyverse"
author: "Grant McDermott, adapted by Kyle Coombs"
date: "Bates College | [EC/DCS 368](https://github.com/ECON368-fall2023-big-data-and-economics)" #"`r format(Sys.time(), '%d %B %Y')`"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
---
name: toc

```{css, echo=FALSE}
@media print {
  .has-continuation {
    display: block !important;
  }
}
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center", #fig.width=6, fig.height=4.5, 
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )
```

# Table of contents

1. [Prologue](#prologue)

2. [Tidyverse basics](#basics)

3. [Data wrangling with dplyr](#dplyr)
  - [filter](#filter)
  - [arrange](#arrange)
  - [select](#select)
  - [mutate](#mutate)
  - [summarise](#summarise)
  - [joins](#joins)

4. [Data tidying with tidyr](#tidyr)
  - [pivot_longer](#pivot_longer) / [pivot_wider](#pivot_wider)
  - [separate](#separate)
  - [unite](#unite)

5. [Summary](#summary)


---
class: inverse, center, middle
name: prologue

# Prologue

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>


---

# Housekeeping 

- Full version of these slides available on [Grant McDermott's website](https://github.com/uo-ec607/lectures)

- Problem Set 2 has been posted and is due 10/9 by Midnight

---

# What is "tidy" data?

### Resources:
- [Vignette](https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html) (from the **tidyr** package)
- [Original paper](https://vita.had.co.nz/papers/tidy-data.pdf) (Hadley Wickham, 2014 JSS)

--

</br>

### Key points:
1. Each variable forms a column.
2. Each observation forms a row.
3. Each type of observational unit forms a table.

--

</br>

Basically, tidy data is more likely to be [long (i.e. narrow) format](https://en.wikipedia.org/wiki/Wide_and_narrow_data) than wide format.

---

# Checklist

### R packages you'll need today

☑ [**tidyverse**](https://www.tidyverse.org/)

☑ [**nycflights13**](hhttps://github.com/hadley/nycflights13)

☑ `data.table`

☑ `tidyfast`

☑ `dtplyr`

☑ `microbenchmark`

--

</br>

The following code chunk will install (if necessary) and load everything for you.

```{r libs, cache=FALSE, message=FALSE}
if (!require(pacman)) install.packages('pacman', repos = 'https://cran.rstudio.com')
pacman::p_load(tidyverse, data.table, dtplyr, tidyfast, microbenchmark, ggplot2, nycflights13)
```

**Tip:** If you're on Linux, then I _strongly_ recommend installing the pre-compiled binary versions of these packages from [RSPM](https://packagemanager.rstudio.com/client/#/repos/1/overview) instead of CRAN. The exact repo mirror varies by distro (see the link). But on Ubuntu 20.04, for example, you'd use: 

```r
install.packages(c('tidyverse', 'nycflights13','data.table','dtplyr','tidyfast','microbenchmark','ggplot2'), repos = 'https://packagemanager.rstudio.com/all/__linux__/focal/latest', dependencies  = TRUE)
library(tidyverse, data.table, dtplyr, tidyfast, microbenchmark, ggplot2, nycflights13)
```
```


---
class: inverse, center, middle
name: basics

# Tidyverse basics

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>

---

# Tidyverse vs. base R

Much digital ink has been spilled over the "tidyverse vs. base R" debate.

--

I won't delve into this debate here, because I think the answer is [clear](http://varianceexplained.org/r/teach-tidyverse/): We should teach the tidyverse first (or, at least, early).
- The documentation and community support are outstanding.
- Having a consistent philosophy and syntax makes it easier to learn.
- Provides a convenient "front-end" to big data tools that we'll use later in the course.
- For data cleaning, wrangling, and plotting, the tidyverse really is a no-brainer.<sup>1</sup>

.footnote[
<sup>1</sup> I'm also a huge fan of [**data.table**](http://r-datatable.com/). This package will be the subject of our next lecture.
]

--

**But**... this certainly shouldn't put you off learning base R alternatives.
- Base R is extremely flexible and powerful (and stable).
- There are some things that you'll have to venture outside of the tidyverse for.
- A combination of tidyverse and base R is often the best solution to a problem.
- Excellent base R data manipulation tutorials: [here](https://www.rspatial.org/intr/index.html) and [here](https://github.com/matloff/fasteR).

---

# Tidyverse vs. base R (cont.)

One point of convenience is that there is often a direct correspondence between a tidyverse command and its base R equivalent. 

These generally follow a `tidyverse::snake_case` vs `base::period.case` rule. E.g. Compare:

| tidyverse  |  base |
|---|---|
| `?readr::read_csv`  | `?utils::read.csv` |
|  `?dplyr::if_else` |  `?base::ifelse` |
|  `?tibble::tibble` |  `?base::data.frame` |
Etcetera.
  
If you call up the above examples, you'll see that the tidyverse alternative typically offers some enhancements or other useful options (and sometimes restrictions) over its base counterpart.

--

**Remember:** There are (almost) always multiple ways to achieve a single goal in R.

---

# Tidyverse packages

Let's load the tidyverse meta-package and check the output.
```{r tverse, cache = FALSE}
library(tidyverse)
```

--

We see that we have actually loaded a number of packages (which could also be loaded individually): **ggplot2**, **tibble**, **dplyr**, etc.
- We can also see information about the package versions and some [namespace conflicts](https://raw.githack.com/uo-ec607/lectures/master/04-rlang/04-rlang.html#59).

---

# Tidyverse packages (cont.)

The tidyverse actually comes with a lot more packages than those that are just loaded automatically.<sup>1</sup>
```{r tverse_pkgs}
tidyverse_packages()
```

We'll use several of these additional packages during the remainder of this course.

- E.g. The **lubridate** package for working with dates and the **rvest** package for webscraping.
- However, bear in mind that these packages will have to be loaded separately.

.footnote[
<sup>1</sup> It also includes a *lot* of dependencies upon installation. This is a matter of some [controversy](http://www.tinyverse.org/).
]

---

# Tidyverse packages (cont.)

I hope to cover most of the tidyverse packages over the length of this course.

Today, however, I'm only really going to focus on two packages: 
1. [**dplyr**](https://dplyr.tidyverse.org/)
2. [**tidyr**](https://tidyr.tidyverse.org/)

These are the workhorse packages for cleaning and wrangling data. They are thus the ones that you will likely make the most use of (alongside **ggplot2**, which we already met back in Lecture 1).
- Data cleaning and wrangling occupies an inordinate amount of time, no matter where you are in your research career.

---
class: inverse, center, middle
name: dplyr

# dplyr

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>

--

*Note:* **dplyr** 1.0.0 also notifies you about grouping variables every time you do operations on or with them. YMMV, but, personally, I find these messages annoying and so prefer to [switch them off](https://twitter.com/MattCowgill/status/1278463099272491008).

```r
options(dplyr.summarise.inform = FALSE) ## Add to .Rprofile to make permanent
```

---

# Key dplyr verbs

There are five key dplyr verbs that you need to learn.

1. `filter`: Filter (i.e. subset) rows based on their values.

2. `arrange`: Arrange (i.e. reorder) rows based on their values.

3. `select`: Select (i.e. subset) columns by their names: 

4. `mutate`: Create new columns.

5. `summarise`: Collapse multiple rows into a single summary value.<sup>1</sup>

.footnote[
<sup>1</sup> `summarize` with a "z" works too, but Hadley Wickham is from New Sealand.
]

--

</br>

Let's practice these commands together using the `starwars` data frame that comes pre-packaged with dplyr. 

---
name: filter

# 1) dplyr::filter

We can chain multiple filter commands with the pipe (`%>%`)[<sup>1<\sup>], or just separate them within a single filter command using commas.
```{r filter1}
starwars %>% 
  filter( 
    species == "Human", 
    height >= 190
    ) 
```

.footnote[<sup>1<\sup> Pipes were invented by Doug McIlroy in 1964, are widely used in Unix shells (e.g. bash) and other programming languages (e.g. `F#`). They pass the preceding object as the first argument to the following function. In R, they
allow you to chain together code in a way that reads from left to right.]


---

# 1) dplyr::filter *cont.*

Regular expressions work well too.
```{r filter2}
starwars %>% 
  filter(grepl("Skywalker", name))
```

---

# 1) dplyr::filter *cont.*

A very common `filter` use case is identifying (or removing) missing data cases. 
```{r filter3}
starwars %>% 
  filter(is.na(height))
```

--

</br>

To remove missing observations, simply use negation: `filter(!is.na(height))`. Try this yourself.

---
name: arrange

# 2) dplyr::arrange

```{r arrange1}
starwars %>% 
  arrange(birth_year)
```

--

*Note:* Arranging on a character-based column (i.e. strings) will sort alphabetically. Try this yourself by arranging according to the "name" column.

---

# 2) dplyr::arrange *cont.*

We can also arrange items in descending order using `arrange(desc())`.
```{r arrange2}
starwars %>% 
  arrange(desc(birth_year))
```

---
name: select

# 3) dplyr::select

Use commas to select multiple columns out of a data frame. (You can also use "first:last" for consecutive columns). Deselect a column with "-".
```{r select1}
starwars %>% 
  select(name:skin_color, species, -height)
```

---

# 3) dplyr::select *cont.*

You can also rename some (or all) of your selected variables in place.
```{r select2}
starwars %>%
  select(alias=name, crib=homeworld, sex=gender) 
```

--

If you just want to rename columns without subsetting them, you can use `rename`. Try this now by replacing `select(...)` in the above code chunk with `rename(...)`.

---

# 3) dplyr::select *cont.*

The `select(contains(PATTERN))` option provides a nice shortcut in relevant cases.
```{r select3}
starwars %>% 
  select(name, contains("color"))
```

---

---
name: mutate

# 4) dplyr::mutate

You can create new columns from scratch, or (more commonly) as transformations of existing columns.
```{r mutate1}
starwars %>% 
  select(name, birth_year) %>%
  mutate(dog_years = birth_year * 7) %>%
  mutate(comment = paste0(name, " is ", dog_years, " in dog years."))
```

---


# 4) dplyr::mutate *cont.*

*Note:* `mutate` is order aware. So you can chain multiple mutates in a single call.
```{r mutate2}
starwars %>% 
  select(name, birth_year) %>%
  mutate(
    dog_years = birth_year * 7, ## Separate with a comma
    comment = paste0(name, " is ", dog_years, " in dog years.")
    )
```

---

# 4) dplyr::mutate *cont.*

Boolean, logical and conditional operators all work well with `mutate` too.
```{r mutate3}
starwars %>% 
  select(name, height) %>%
  filter(name %in% c("Luke Skywalker", "Anakin Skywalker")) %>% 
  mutate(tall1 = height > 180) %>%
  mutate(tall2 = ifelse(height > 180, "Tall", "Short")) ## Same effect, but can choose labels

```

---

# 4) dplyr::mutate *cont.*

Lastly, combining `mutate` with the `across` feature allows you to easily work on a subset of variables. For example:

```{r, mutate4}
starwars %>% 
  select(name:eye_color) %>% 
  mutate(across(where(is.character), toupper)) %>% #<< 
  head(5)
```

---
name: summarise

# 5) dplyr::summarise

Particularly useful in combination with the `group_by` command.
```{r summ1}
starwars %>% 
  group_by(species, gender) %>% 
  summarise(mean_height = mean(height, na.rm = TRUE))
```

---

# 5) dplyr::summarise *cont.*

Note that including "na.rm = TRUE" (or,  its alias "na.rm = T") is usually a good idea with summarise functions. Otherwise, your output will be missing too. 
```{r summ2}
## Probably not what we want
starwars %>% 
  summarise(mean_height = mean(height))
## Much better
starwars %>% 
  summarise(mean_height = mean(height, na.rm = TRUE))
```

---

# 5) dplyr::summarise *cont.*

The same `across`-based workflow that we saw with `mutate` a few slides back also works with `summarise`. For example:

```{r, summ4}
starwars %>% 
  group_by(species) %>% 
  summarise(across(where(is.numeric), mean, na.rm=T)) %>% #<<
  head(5)
```

---

# Other dplyr goodies

`group_by` and `ungroup`: For (un)grouping.
- Particularly useful with the `summarise` and `mutate` commands, as we've already seen.

--

`slice`: Subset rows by position rather than filtering by values.
- E.g. `starwars %>% slice(c(1, 5))`

--

`pull`: Extract a column from as a data frame as a vector or scalar.
- E.g. `starwars %>% filter(gender=="female") %>% pull(height)`

--

`count` and `distinct`: Number and isolate unique observations.
- E.g. `starwars %>% count(species)`, or `starwars %>% distinct(species)`
- You could also use a combination of `mutate`, `group_by`, and `n()`, e.g. `starwars %>% group_by(species) %>% mutate(num = n())`.


---

# Other dplyr goodies (cont.)

There are also a whole class of [window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html) for getting leads and lags, ranking, creating cumulative aggregates, etc.
- See `vignette("window-functions")`.

--

</br>

The final set of dplyr "goodies" are the family of join operations. However, these are important enough that I want to go over some concepts in a bit more depth...
- We will encounter and practice these many more times as the course progresses.

---
name: joins

# Joins

One of the mainstays of the dplyr package is merging data with the family [join operations](https://cran.r-project.org/web/packages/dplyr/vignettes/two-table.html).
- `inner_join(df1, df2)`
- `left_join(df1, df2)`
- `right_join(df1, df2)`
- `full_join(df1, df2)`
- `semi_join(df1, df2)`
- `anti_join(df1, df2)`

Joins are how you get .hi[Relational Database Managment] (RDBM) to work in R. 

(See visual depictions of the different join operations [here](https://r4ds.had.co.nz/relational-data.html).)


--

For the simple examples that I'm going to show here, we'll need some data sets that come bundled with the [**nycflights13**](http://github.com/hadley/nycflights13) package. 
- Load it now and then inspect these data frames in your own console.

```{r flights, echo = F}
library(nycflights13)
```
```{r, eval = F}
library(nycflights13)
flights 
planes
```

- Any variable names in common?
- Are the definitions of these variables consistent across data frames?

---

# Joins (cont.)

Let's perform a [left join](https://stat545.com/bit001_dplyr-cheatsheet.html#left_joinsuperheroes-publishers) on the flights and planes datasets. 
- *Note*: I'm going subset columns after the join, but only to keep text on the slide.

--

```{r join1}
left_join(flights, planes) %>%
  select(year, month, day, dep_time, arr_time, carrier, flight, tailnum, type, model)
```

---

# Joins (cont.)

(*continued from previous slide*)

Note that dplyr made a reasonable guess about which columns to join on (i.e. columns that share the same name). It also told us its choices: 

```
*## Joining, by = c("year", "tailnum")
```

However, there's a problem here: the variable "year" does not have a consistent meaning across our joining datasets!
- In one it refers to the *year of flight*, in the other it refers to *year of construction*.

--

Luckily, there's an easy way to avoid this problem. 
- See if you can figure it out before turning to the next slide.
- Try `?dplyr::join`.

---

# Joins (cont.)

(*continued from previous slide*)

You just need to be more explicit in your join call by using the `by = ` argument.
- You can also rename any ambiguous columns to avoid confusion. 
```{r join2}
left_join(
  flights,
  planes %>% rename(year_built = year), ## Not necessary w/ below line, but helpful
  by = "tailnum" ## Be specific about the joining column
  ) %>%
  select(year, month, day, dep_time, arr_time, carrier, flight, tailnum, year_built, type, model) %>%
  head(3) ## Just to save vertical space on the slide
```

---

# Joins (cont.)

(*continued from previous slide*)

Last thing I'll mention for now; note what happens if we again specify the join column... but don't rename the ambiguous "year" column in at least one of the given data frames.
```{r join3}
left_join(
  flights,
  planes, ## Not renaming "year" to "year_built" this time
  by = "tailnum"
  ) %>%
  select(contains("year"), month, day, dep_time, arr_time, carrier, flight, tailnum, type, model) %>%
  head(3)
```

--

Make sure you know what "year.x" and "year.y" are. Again, it pays to be specific.

---
class: inverse, center, middle
name: tidyr

# tidyr

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>

---

# Key tidyr verbs

1. `pivot_longer`: Pivot wide data into long format (i.e. "melt").<sup>1</sup> 

2. `pivot_wider`: Pivot long data into wide format (i.e. "cast").<sup>2</sup> 

3. `separate`: Separate (i.e. split) one column into multiple columns.

4. `unite`: Unite (i.e. combine) multiple columns into one.

.footnote[
<sup>1</sup> Updated version of `tidyr::gather`.

<sup>2</sup> Updated version of `tidyr::spread`.
]  

--

</br>

Let's practice these verbs together in class.
- Side question: Which of `pivot_longer` vs `pivot_wider` produces "tidy" data?
  
---
name: pivot_longer

# 1) tidyr::pivot_longer

```{r pivot_longer1}
stocks = data.frame( ## Could use "tibble" instead of "data.frame" if you prefer
  time = as.Date('2009-01-01') + 0:1,
  X = rnorm(2, 0, 1),
  Y = rnorm(2, 0, 2),
  Z = rnorm(2, 0, 4)
  )
stocks
stocks %>% pivot_longer(-time, names_to="stock", values_to="price")
```

---

# 1) tidyr::pivot_longer *cont.*

Let's quickly save the "tidy" (i.e. long) stocks data frame for use on the next slide. 

```{r pivot_longer2}
## Write out the argument names this time: i.e. "names_to=" and "values_to="
tidy_stocks = 
  stocks %>% 
  pivot_longer(-time, names_to="stock", values_to="price")
```

---
name: pivot_wider

# 2) tidyr::pivot_wider

```{r pivot_wider1}
tidy_stocks %>% pivot_wider(names_from=stock, values_from=price)
tidy_stocks %>% pivot_wider(names_from=time, values_from=price)
```

--

</br>
Note that the second example &mdash; which has combined different pivoting arguments &mdash; has effectively transposed the data.


---

# Aside: Remembering the pivot_* syntax 

There's a long-running joke about no-one being able to remember Stata's "reshape" command. ([Exhibit A](https://twitter.com/helleringer143/status/1117234887902285836).)

It's easy to see this happening with the `pivot_*` functions too. However, I find that I never forget the commands as long as I remember the argument order is *"names"* then *"values"*.

---
name: separate

# 3) tidyr::separate

```{r sep1}
economists = data.frame(name = c("Adam.Smith", "Paul.Samuelson", "Milton.Friedman"))
economists
economists %>% separate(name, c("first_name", "last_name")) 
```

--

</br>

This command is pretty smart. But to avoid ambiguity, you can also specify the separation character with `separate(..., sep=".")`.

---

# 3) tidyr::separate *cont.*

A related function is `separate_rows`, for splitting up cells that contain multiple fields or observations (a frustratingly common occurence with survey data).

```{r sep2}
jobs = data.frame(
  name = c("Jack", "Jill"),
  occupation = c("Homemaker", "Philosopher, Philanthropist, Troublemaker") 
  ) 
jobs
## Now split out Jill's various occupations into different rows
jobs %>% separate_rows(occupation)
```

---
name: unite

# 4) tidyr::unite

```{r unite1}
gdp = data.frame(
  yr = rep(2016, times = 4),
  mnth = rep(1, times = 4),
  dy = 1:4,
  gdp = rnorm(4, mean = 100, sd = 2)
  )
gdp 
## Combine "yr", "mnth", and "dy" into one "date" column
gdp %>% unite(date, c("yr", "mnth", "dy"), sep = "-")
```

---

# 4) tidyr::unite *cont.*

Note that `unite` will automatically create a character variable. You can see this better if we convert it to a tibble. 
```{r unite2}
gdp_u = gdp %>% unite(date, c("yr", "mnth", "dy"), sep = "-") %>% as_tibble()
gdp_u
```

--

If you want to convert it to something else (e.g. date or numeric) then you will need to modify it using `mutate`. See the next slide for an example, using the [lubridate](https://lubridate.tidyverse.org/) package's super helpful date conversion functions.

---

# 4) tidyr::unite *cont.*

*(continued from previous slide)*

```{r unite3, message=F}
library(lubridate)
gdp_u %>% mutate(date = ymd(date))
```

---
class: inverse, center, middle
name: summary

# Summary
<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>

---

# Key verbs

### dplyr
1. `filter`
2. `arrange`
3. `select`
4. `mutate`
5. `summarise`

### tidyr
1. `pivot_longer`
2. `pivot_wider`
3. `separate`
4. `unite`

--

Other useful items include: pipes (`%>%`), grouping (`group_by`), joining functions (`left_join`, `inner_join`, etc.).

---
class: inverse, center, middle

# Next lecture: Data cleaning and wrangling: (2) data.table
<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>

```{r gen_pdf, include = FALSE, cache = FALSE, eval = TRUE}
infile = list.files(pattern = '.html')
pagedown::chrome_print(input = infile, timeout = 100)
```