---
title: Big Data and Economics
subtitle: Text as Data: Regular Expressions and Wordclouds
author:
  name: Kyle Coombs
  affiliation: Bates College | [DCS/ECON 368](https://github.com/ECON368-fall2023-big-data-and-economics/big-data-class-materials)  
output:
  html_document:
    theme: journal
    highlight: haddock
    # code_folding: show
    toc: yes
    toc_depth: 4
    toc_float: yes
    keep_md: true
    keep_tex: false ## Change to true if want keep intermediate .tex file
    css: css/preamble.css ## For multi-col environments
  pdf_document:
    latex_engine: xelatex
    toc: true
    dev: cairo_pdf
    # fig_width: 7 ## Optional: Set default PDF figure width
    # fig_height: 6 ## Optional: Set default PDF figure height
    includes:
      in_header: tex/preamble.tex ## For multi-col environments
    extra_dependencies: [float, booktabs, longtable]
    pandoc_args:
        --template=tex/mytemplate.tex ## For affiliation field. See: https://bit.ly/2T191uZ
always_allow_html: true
urlcolor: blue
mainfont: cochineal
sansfont: Fira Sans
#monofont: Fira Code ## Although, see: https://tex.stackexchange.com/q/294362
## Automatically knit to both formats:
knit: (function(inputFile, encoding) {
 rmarkdown::render(inputFile, encoding = encoding, 
 output_format = 'all') 
 })
---

```{r setup, include=FALSE}
knitr::opts_chunk(echo = TRUE, cache = TRUE, dpi=300)
```


### Fixed effects using a familiar dataset 

Let's review fixed effects very quickly using the Ask A Manager Survey 2023, which you used on problem set 2. Like on your problem set, we'll load it in using the **gsheet** package. Several of you said that you thought you could use fixed effects to residualize other differences between groups out of the data. Let's see if that's true using the package **fixest**. _As a disclaimer: these data are not systematically collected._

```{r manager}
column_names <- c('timestamp','age','industry','area','jobtitle','jobtitle2',
        'annual_salary','additional_pay','currency','currency_other',
        'income_additional','country','state','city','remote','experience_overall',
        'experience_field','education','gender','race')

US_strings<-c("United States of America","United States", 
  "United states" ,"USA","Usa","usa" ,"US","U.S." ,"us")

managers = gsheet::gsheet2tbl(
    'https://docs.google.com/spreadsheets/d/1ioUjhnz6ywSpEbARI-G3RoPyO0NRBqrJnWf-7C_eirs/edit?resourcekey#gid=1854892322') 
names(managers) <- column_names

# Focus on the US and cismen and women to simplify the analysis
managers <- managers %>%
  filter(country %in% US_strings,
    gender %in% c('Man','Woman'))
```

First, let's measure the gender pay gap as the average difference in pay. 

```{r gender-pay-gap}
pay_gap <- managers %>% 
  group_by(gender) %>%
  summarize(mean_salary = mean(annual_salary, na.rm = TRUE)) %>%
  ungroup() %>%
  summarize(gender_pay_gap = diff(mean_salary))

print(paste('The average difference in pay is', pay_gap$gender_pay_gap))
```

What would the coefficient on gender be in this regression? 

$$ 
\begin{align}
\text{Salary}_{i} ~ \beta_0 + \beta_1 \text{Gender}_i + \varepsilon_{i}
\end{align}
$$

where Gender takes a value of 1 if the person is a woman. Make a guess before you estimate below. 

.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
```{r gender-dummy}
lm(annual_salary ~ gender, data=managers)
```

That's right, it is the average difference in pay. Why is that? Well what is the intercept from the regression? 

The average salary for men. So when $Gender_i=0$, the regression returns the average salary for men because this is the "Best Linear Unbiased Estimate" of salary, which is the goal of OLS. When $Gender_i=1$, OLS returns the average salary for women, again for the same reason! 

That dummy variable is actually a fixed effect for gender. We could add in many, this is called the Least Squares Dummy Variable approach.^[This name comes from Nick Huntington-Klein.] Lets add a dummy for remote work.

```{r other-dummies}
lm(annual_salary ~ gender + remote, data=managers)
```

Do you see how each level of remote got its own bracket? That's because each level of remote is a fixed effect. Do you also see how the 