---
title: "Midterm project"
subtitle: "EC 421: Introduction to Econometrics"
# author: "Edward Rubin"
date: "Due *before* midnight on Sunday, 21 February 2021"
# date: ".it.biggest[Solutions]"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    # self_contained: true
    nature:
      ratio: '8.5:11'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
layout: true
class: clear
---

```{r, setup, include = F}
# Knitr options
library(knitr)
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  warning = F,
  message = F
)
opts_chunk$set(dev = "svg")
options(device = function(file, width, height) {
  svg(tempfile(), width = width, height = height)
})
options(digits = 4)
options(width = 90)
```

## Instructions

.mono.b[INTEGRITY:] Groups can either have **one or two members**. Only one person needs to submit your final document. If you are suspected of cheating in any way (for example, copying from someone else), then you will receive a zero and fail this course. We will report you to the dean.

.mono.b[GRADING:] Your grade for this project will be based upon the accuracy of your answers *and* how well you explain/illustrate your answers. We value short, accurate answers over long, meandering answers. Edit your answers! Make your figures look good (including titles and labeled axes)!

.mono.b[EMAIL POLICY:] Do not ask the GEs, the instructor, or people outside your group for help coding or for help answering these questions. You may only ask **clarifying** questions. **Use Google and the course's materials** (lectures, labs, notes, assignment keys).

.mono.b[DUE:] **One member** of your group must upload your answer on [Canvas](https://canvas.uoregon.edu/) *before* midnight on Sunday, 21 February 2021. All members of the group must be listed on the submission.

.mono.b[IMPORTANT:] As with your homework, you must submit .b[two files]:
<br> .b.mono[1.] your typed responses/answers to the question **with figures and regression results** (in a Word file or something similar).
<br> .b.mono[2.] the .mono[R] script you used to generate your answers. 

If you are using RMarkdown, you can submit a single file.

.mono.b[README!] The last page has a table that describes each variable in the dataset (`data-project-01.csv`).

.mono.b[HELP!] The questions below ask for several figures. If you need help creating the figures, check out these `ggplot2` resources (in addition to the class and lab materials):

- [An intro to `ggplot2`](https://uc-r.github.io/ggplot_intro)
- [A tutorial on customizing `ggplot2` figures](http://r-statistics.co/Complete-Ggplot2-Tutorial-Part2-Customizing-Theme-With-R-Code.html)
- [The `ggplot2` website (with a `ggplot2` cheatsheet!)](https://ggplot2.tidyverse.org/index.html)

---

## Questions

.b[01.] Load the data (`data-project-01.csv`). Summarize and describe the variables in the dataset. Your answer should include:

- Which countries show up in the data? What are their percentages (share of the sample)?
- How many provinces and "tasters" show up in the data? How many varieties are there?
- How skewed are the distributions of price and points?
- Create at least two figures (graphs) that individually summarize the variables `price` and `points`.
- Create at least three figures (graphs) that demonstrate how the key variables relate to each other and other variables (*i.e.*, `variety`, `country`, `province`, `taster_name`).

Explain your decisions on summarizing the data. What do you learn about potential relationships?

.b[02.] Does the distribution of `price` appear to be the same across the countries? What about the distribution of `points` across the countries? Use figures (plots) to justify your answer. Explain your answer.

.b[03.] We are going to treat `price` as our outcome variable. Regress the price (of the wine bottle) on an intercept, the bottle's rating (`points`) and its country of origin. Explain what the coefficients mean and comment on their statistical significance.

.b[04.] Create a scatter plot with the residuals from .b[03] on the y axis the rating (`points`) on the x axis.

.b[05.] Does the scatter plot from .b[04] suggest that **heteroskedasticity** may be present? Explain your answer.

.b[06.] Does the scatter plot from .b[04] suggest that there are any issues with **your specification**? Explain.

.b[07.] Do you think your regression in .b[03] could suffer from omitted-variable bias? Explain why or why not, using the requirements for omitted-variable bias as part of your explanation.

.b[08.] Now include an interaction between country and rating (`points`). Interpret this interaction and comment on the statistical significance. Does this interaction seem important? Explain.

.b[09.] Up to this point, we've told you which regressions to run. And we've stuck with pretty simple regressions (*e.g.*, regress `y` on `x1 + x2`). Now we want you to explore the actual complexity of econometric/statistical analyses. **Estimate three new models.** These models **should not match** your previous models (in .b[03] and .b[08]). Be creative!! Across these three new models, you should include (at least once):

- a log-transformed outcome variable (*i.e.*, use `log()`)
- new/additional explanatory variables (you've only used two of the variables in the dataset)
- an interaction

.b[10.] How did you choose your specifications in .b[09]? Explain your decision making.

.b[11.] Which of your new models is the "best"—*i.e.*, if you must choose one model, which would you choose? Why? Explain your reasoning.

.b[12.] For your "best" model (chosen in .b[11]): Interpret the coefficients and comment on their statistical significance.

.b[13.] Do you *trust* the estimates from your *best model*? Explain why/why not.

.b[14.] Create a scatter plot with the actual prices (`price`) on the *y* axis the the predictions (see the `fitted.values` outputted by `lm()` or use the `fitted()` or `predict()` functions) on the *x* axis. How well does your model predict the true price? Explain your answer.

.b[15.] Write up a one-paragraph summary of what you've learned about pricing in wine. Base your findings on the figures and regressions in this project.

---

```{r, background variables, echo = F, message = F}
# Load requisite packages
pacman::p_load(tidyverse, knitr, kableExtra, here)
# Load data
proj_df = here("data-project-01.csv") %>% read_csv()
# Create table of variable descriptions
var_tbl = data.frame(
  Variable = names(proj_df) %>% paste0(".mono-small[", ., "]"),
  Description = c(
  	"The price of the bottle of wine (US dollars).",
  	"The points given in the review (more points means better wine).",
    "The wine's variety (think: type).",
    "The country that produced the wine.",
    "The province that produced the wine.",
    "The name of the person who tasted/reviewed/rated the bottle of wine."
  )
)
kable(var_tbl) %>%
  kable_styling(full_width = F)
```

---
exclude: true

```{r, print pdf, echo = F, eval = F}
pagedown::chrome_print("prompt-01.html", output = "project-01.pdf")
```