---
title: "Further Examination of an Important Research Question"
author: "Student Name"
date: "May 9, 2020"
output:
  bookdown::html_document2: default
  bookdown::pdf_document2: default
bibliography: References.bib
---

```{r setup, include=FALSE}
#install.packages('lorem','pander','fixest','stargazer','tidyverse','knir')
# Also need to install either install.packages('pdftools') 
# Alternatively, can install ImageMagick and Ghostscript for the fixest package to make pretty tables
library(knitr)
library(lorem)
library(stargazer)
library(fixest)
library(pander)
library(tidyverse)
dict = c("(Intercept)" = "Constant",
         'x1' = "Variable 1",
         'y' = "Outcome",
         'treat' = "Treatment",
        note1 = dsb("*Notes*: This is a note that illustrates how to access notes ",
                    "from the dictionary."),
        source = "*Sources*: Somewhere from the net.")
setFixest_dict(dict)
# The style of the table
my_style = style.tex("aer", model.format = "(i)")
#markdown = TRUE is only useful in Rmarkdown documents
setFixest_etable(style.tex = my_style, markdown = TRUE, page.width = "a4")

```


# Introduction

`r lorem::ipsum(paragraphs = 4)`

# Literature Review

Previous work by @altonji1993 shows that educational decisions are an important determinant of later-life earnings. This point is driven further in follow-up work by @altonji_al2012 and @altonji_al2016.

# Data 

`r lorem::ipsum(paragraphs = 5)`

# Empirical Methods

While my approach explores a number of different approaches, the primary empirical model can be depicted in the following equation:

\[
Y_{it} = \alpha_{0} + \alpha_{1}Z_{it} + \alpha_{2} X_{it} + \varepsilon (\#eq:mean),
\]

where \(Y_{it}\) is a continuous outcome variable for unit \(i\) in year \(t\), and \(Z_{it}\) are characteristics about the firm at which \(i\) is working, while \(X_{it}\) are characteristics about \(i\). The parameter of interest is \(\alpha_{1}\).

# Research Findings

The main results are reported in Table 1. Also in figure \@ref(fig:myfigure). Note the table reference does not work. Needs some more troubleshooting. 

# Conclusion

`r lorem::ipsum(paragraphs = 3)`

# References

<div id="refs"></div>

# Tables 

```{r,results='asis',echo=FALSE}
# Load required packages
# Generate sample data and perform regression
data(mtcars)
lm_model <-  lm(mpg ~ wt, data = mtcars)
lm_model2 <- lm(mpg ~ wt + vs, data = mtcars)
lm_model3 <- lm(mpg ~ wt + vs + as.factor(gear), data = mtcars)

# Generate the regression table
stargazer(lm_model, lm_model2, lm_model3, 
          title = "Linear Regression Results", 
          align = TRUE, 
          style = "qje",
          column.labels = c("Outcome", "Dependent Variable", "Coefficient (SE)"),
          dep.var.caption = "Dependent variable:",
          dep.var.labels = "MPG",
          covariate.labels=c('Weight', 'Straight-Engine', 'Gears 3', 'Gears 4'),
          ci = TRUE,
          single.row = TRUE,
          type='html')

```

Or a table using the fixest, etable functionality. This may be a little trickier to implement. Follow [vignette](https://lrberge.github.io/fixest/articles/etable_new_features.html) here.

```{r ols, echo=FALSE,results='asis'}
data(base_did)
ols = feols(y ~ x1, base_did)
ols_fe = feols(y ~ x1 | id+period, base_did)
iv = feols(y ~ 1 | x1~treat, base_did)
iv_fe = feols(y ~ 1| period | x1 ~ treat, base_did)

etable(list('OLS'=ols, 'OLS'=ols_fe, 
    'First-stage'=summary(iv, stage = 1), 'IV'=summary(iv, stage = 2), 
    'First-stage'=summary(iv_fe, stage = 1:2), 'IV'=summary(iv_fe, stage = 2)),
    style.df = style.df(depvar.title = "", yesNo = "Yes"),
    notes = c("note1", "source"))
```

# Figures

```{r myfigure, fig.cap='My Figure', echo=FALSE}
base_did %>% group_by(treat,period) %>% summarise(y=mean(y)) %>% ungroup() %>%
ggplot(aes(x=period,y=y,col=factor(treat))) + 
         geom_line() +
         labs(y='Outcome',x='Period',col='Treatment',title='Very cool graph of main descriptive result') +
         theme_bw() + 
         theme(legend.position = c(.1,.8),
             text = element_text(size=20)) #bigger font size 
```