---
title: "Further Examination of an Important Research Question"
author: "Student Name"
date: "May 9, 2020"
output:
  bookdown::html_document2: default
  #bookdown::pdf_document2: default
bibliography: References.bib
---

```{r setup, include=FALSE}
#install.packages("lorem",'pander','kableExtra')
library(knitr)
library(kableExtra)
library(xtable)
library(lorem)
library(fixest)
library(pander)
library(tidyverse)
dict = c("(Intercept)" = "Constant", 
         note1 = dsb("*Notes*: This is a note that illustrates how to access notes ",
                     "from the dictionary."),
         source = "*Sources*: Somewhere from the net.")
setFixest_dict(dict)
# The style of the table

# markdown = TRUE is only useful in Rmarkdown documents
```

# Introduction

`r lorem::ipsum(paragraphs = 4)`

# Literature Review

Previous work by @altonji1993 shows that educational decisions are an important determinant of later-life earnings. This point is driven further in follow-up work by @altonji_al2012 and @altonji_al2016.

# Data 

`r lorem::ipsum(paragraphs = 5)`

# Empirical Methods

While my approach explores a number of different approaches, the primary empirical model can be depicted in the following equation:

\[
Y_{it} = \alpha_{0} + \alpha_{1}Z_{it} + \alpha_{2} X_{it} + \varepsilon (\#eq:mean),
\]

where \(Y_{it}\) is a continuous outcome variable for unit \(i\) in year \(t\), and \(Z_{it}\) are characteristics about the firm at which \(i\) is working, while \(X_{it}\) are characteristics about \(i\). The parameter of interest is \(\alpha_{1}\).

# Research Findings

The main results are reported in Table \@ref(mytable). Also in figure \@ref(fig:myfigure). Note the table reference does not work. Needs some more troubleshooting. 

# Conclusion

`r lorem::ipsum(paragraphs = 3)`

# References

<div id="refs"></div>

# Tables 

 ```{r ols, echo=FALSE}
data(trade)
data(base_did)
ols = feols(y ~ x1, base_did)
ols_fe = feols(y ~ x1 | id+period, base_did)
iv = feols(y ~ 1 | x1~treat, base_did)
iv_fe = feols(y ~ 1| period | x1 ~ treat, base_did)

# Create tables using kable
```


```{r mytable,echo=FALSE}
#Knitr::kable allows the cross-reference of tables, etable makes a pretty table.
etable(list('OLS'=ols, 'OLS'=ols_fe, 
    'First-stage'=summary(iv, stage = 1), 'IV'=summary(iv, stage = 2), 
    'First-stage'=summary(iv_fe, stage = 1:2), 'IV'=summary(iv_fe, stage = 2)), 
    style.df = style.df(depvar.title = "", 
    yesNo = "yes"),
    notes = c("note1", "source"),
    label='mytable',
    postprocess.df = pandoc.table.return, style = "rmarkdown")
```

```{r myfigure, fig.cap='My Figure'}
base_did %>% group_by(treat,period) %>% summarise(y=mean(y)) %>%
    ggplot(aes(x=period,y=y,col=factor(treat))) + 
        geom_line() +
        labs(y='Outcome',x='Period',col='Treatment',title='Very cool graph of main descriptive result') +
        theme_bw() +
        theme(legend.position = c(.1,.8),
            text = element_text(size=20),) #bigger font size

```