---
title: Data Science for Economists
subtitle: Application of LASSO and Ridge
author:
  name: Kyle Coombs (adapted from [ISLR tidymodels labs](https://emilhvitfeldt.github.io/ISLR-tidymodels-labs/06-regularization.html))
  affiliation: Bates College | [DCS/ECON 368](https://github.com/big-data-and-economics/big-data-class-materials)  
output:
  html_document:
    theme: journal
    highlight: haddock
    # code_folding: show
    toc: yes
    toc_depth: 4
    toc_float: yes
    keep_md: true
    keep_tex: false ## Change to true if want keep intermediate .tex file
    css: css/preamble.css ## For multi-col environments
  pdf_document:
    latex_engine: xelatex
    toc: true
    dev: cairo_pdf
    # fig_width: 7 ## Optional: Set default PDF figure width
    # fig_height: 6 ## Optional: Set default PDF figure height
    includes:
      in_header: tex/preamble.tex ## For multi-col environments
    extra_dependencies: [float, booktabs, longtable]
    pandoc_args:
        --template=tex/mytemplate.tex ## For affiliation field. See: https://bit.ly/2T191uZ
always_allow_html: true
urlcolor: blue
mainfont: cochineal
sansfont: Fira Sans
#monofont: Fira Code ## Although, see: https://tex.stackexchange.com/q/294362
## Automatically knit to both formats:
knit: (function(inputFile, encoding) {
 rmarkdown::render(inputFile, encoding = encoding, 
 output_format = 'all') 
 })
---

You need the following packages to practice this cute little vignette based off of work by Emil Hvitfeldt on labs in the book Introduction to Statistical Learning with R using the R package **tidymodels**. Please answer the following questions and push your answers to your GitHub repository for class.

1. What are the coefficients in the Ridge and LASSO regressions when the penalty is zero? Why? 
2. How does tidymodels pick the optimal $\lambda$ in each method? 
3. What is the optimal $\lambda$ in Ridge and LASSO?

```{r libs_print, cache=FALSE, eval=FALSE}
## Load and install the packages that we'll be using today
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidymodels,ISLR,ggplot,glmnet,tidyverse)

## My preferred ggplot2 plotting theme (optional)
theme_set(theme_minimal())
```
```{r libs, cache=FALSE, message=FALSE, echo=FALSE}
## Load and install the packages that we'll be using today
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidymodels,ISLR,ggplot,glmnet,tidyverse)## My preferred ggplot2 plotting theme (optional)
theme_set(theme_minimal())

## Set future::plan() resolution strategy
```

# Linear Model Selection and Regularization

This lab will take a look at regularization models and hyperparameter tuning. These models are related to the models we saw in chapter 3 and 4, with the difference that they contain a regularization term.
This chapter will use [parsnip](https://www.tidymodels.org/start/models/) for model fitting and [recipes and workflows](https://www.tidymodels.org/start/recipes/) to perform the transformations, and [tune and dials](https://www.tidymodels.org/start/tuning/) to tune the hyperparameters of the model.

We will be using the `Hitters` data set from the `ISLR` package. We wish to predict the baseball players `Salary` based on several different characteristics which are included in the data set. Since we wish to predict `Salary`, then we need to remove any missing data from that column. Otherwise, we won't be able to run the models.

```{r}
#| message: false
Hitters <- as_tibble(Hitters) %>%
  filter(!is.na(Salary))
```


## Ridge Regression

We will use the `glmnet` package to perform ridge regression. `parsnip` does not have a dedicated function to create a ridge regression model specification. You need to use `linear_reg()` and set `mixture = 0` to specify a ridge model. The `mixture` argument specifies the amount of different types of regularization, `mixture = 0` specifies only ridge regularization and `mixture = 1` specifies only lasso regularization. Setting `mixture` to a value between 0 and 1 lets us use both. When using the `glmnet` engine we also need to set a `penalty` to be able to fit the model. We will set this value to `0` for now, it is not the best value, but we will look at how to select the best value in a little bit.

```{r}
ridge_spec <- linear_reg(mixture = 0, penalty = 0) %>%
  set_mode("regression") %>%
  set_engine("glmnet")
```

Once the specification is created we can fit it to our data. We will use all the predictors.

```{r}
ridge_fit <- fit(ridge_spec, Salary ~ ., data = Hitters)
```

The `glmnet` package will fit the model for all values of `penalty` at once, so let us see what the parameter estimate for the model is now that we have `penalty = 0`.

```{r}
#| message: false
tidy(ridge_fit)
```

Let us instead see what the estimates would be if the penalty was 11498.

```{r}
tidy(ridge_fit, penalty = 11498)
```

Notice how the estimates are decreasing when the amount of penalty goes up. Look below at the parameter estimates for `penalty = 705` and `penalty = 50`.

```{r}
tidy(ridge_fit, penalty = 705)
tidy(ridge_fit, penalty = 50)
```

We can visualize how the magnitude of the coefficients are being regularized towards zero as the penalty goes up. 

```{r}
#| fig-alt: |
#|   Multiple line chart. Log Lambda along the x-axis, Coefficients
#|   along the y-axis. The curves starts at different places along
#|   the y-axis but slowly converge towards 0 as the Log Lambda 
#|   value increase.
ridge_fit %>%
  autoplot()
```

Prediction is done like normal, if we use `predict()` by itself, then `penalty = 0` as we set in the model specification is used.

```{r}
predict(ridge_fit, new_data = Hitters)
```

but we can also get predictions for other values of `penalty` by specifying it in `predict()`

```{r}
predict(ridge_fit, new_data = Hitters, penalty = 500)
```

We saw how we can fit a ridge model and make predictions for different values of `penalty`. But it would be nice if we could find the "best" value of the penalty. This is something we can use hyperparameter tuning for. Hyperparameter tuning is in its simplest form a way of fitting many models with different sets of hyperparameters trying to find one that performs "best". The complexity in hyperparameter tuning can come from how you try different models. We will keep it simple for this lab and only look at grid search, only looking at evenly spaced parameter values. This is a fine enough approach if you have one or two tunable parameters but can become computationally infeasible. See the chapter on [iterative search](https://www.tmwr.org/iterative-search.html) from [Tidy Modeling with R](https://www.tmwr.org/) for more information.

We start like normal by setting up a validation split. A K-fold cross-validation data set is created on the training data set with 10 folds.

```{r}
Hitters_split <- initial_split(Hitters, strata = "Salary")

Hitters_train <- training(Hitters_split)
Hitters_test <- testing(Hitters_split)

Hitters_fold <- vfold_cv(Hitters_train, v = 10)
```

We can use the `tune_grid()` function to perform hyperparameter tuning using a grid search. `tune_grid()` needs 3 different thing;

- a `workflow` object containing the model and preprocessor,
- a `rset` object containing the resamples the `workflow` should be fitted within, and
- a tibble containing the parameter values to be evaluated.

Optionally a metric set of performance metrics can be supplied for evaluation. If you don't set one then a default set of performance metrics is used.

We already have a resample object created in `Hitters_fold`. Now we should create the workflow specification next.

We just used the data set as is when we fit the model earlier. But ridge regression is scale sensitive so we need to make sure that the variables are on the same scale. We can use `step_normalize()`. Secondly let us deal with the factor variables ourself using `step_novel()` and `step_dummy()`.

```{r}
ridge_recipe <- 
  recipe(formula = Salary ~ ., data = Hitters_train) %>% 
  step_novel(all_nominal_predictors()) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_predictors())
```

The model specification will look very similar to what we have seen earlier, but we will set `penalty = tune()`. This tells `tune_grid()` that the `penalty` parameter should be tuned.

```{r}
ridge_spec <- 
  linear_reg(penalty = tune(), mixture = 0) %>% 
  set_mode("regression") %>% 
  set_engine("glmnet")
```

Now we combine to create a `workflow` object.

```{r}
ridge_workflow <- workflow() %>% 
  add_recipe(ridge_recipe) %>% 
  add_model(ridge_spec)
```

The last thing we need is the values of `penalty` we are trying. This can be created using `grid_regular()` which creates a grid of evenly spaces parameter values. We use the `penalty()` function from the [dials](https://dials.tidymodels.org/) package to denote the parameter and set the range of the grid we are searching for. Note that this range is log-scaled.

```{r}
penalty_grid <- grid_regular(penalty(range = c(-5, 5)), levels = 50)
penalty_grid
```

Using 50 levels for one parameter might seem overkill and in many applications it is. But remember that `glmnet` fits all the models in one go so adding more levels to `penalty` doesn't affect the computational speed much.

Now we have everything we need and we can fit all the models.

```{r}
tune_res <- tune_grid(
  ridge_workflow,
  resamples = Hitters_fold, 
  grid = penalty_grid
)

tune_res
```

The output of `tune_grid()` can be hard to read by itself unprocessed. `autoplot()` creates a great visualization 

```{r}
#| fig-alt: |
#|   Facetted connected scatter chart. regularization along the
#|   x-axis. Performance values along the y-axis. The facets are
#|   rmse and rsq. Both are fairly constant for low values of
#|   regularization, rmse starts moderately increasing and rsq
#|   starts moderately decreasing once the regularization
#|   gets larger.
autoplot(tune_res)
```
Here we see that the amount of regularization affects the performance metrics differently. Note how there are areas where the amount of regularization doesn't have any meaningful influence on the coefficient estimates. We can also see the raw metrics that created this chart by calling `collect_matrics()`. 

```{r}
collect_metrics(tune_res)
```

The "best" values of this can be selected using `select_best()`, this function requires you to specify a `matric` that it should select against. 

```{r}
best_penalty <- select_best(tune_res, metric = "rsq")
best_penalty
```

This value of `penalty` can then be used with `finalize_workflow()` to update/finalize the recipe by replacing `tune()` with the value of `best_penalty`. Now, this model should be fit again, this time using the whole training data set.

```{r}
ridge_final <- finalize_workflow(ridge_workflow, best_penalty)

ridge_final_fit <- fit(ridge_final, data = Hitters_train)
```

This final model can now be applied on our testing data set to validate the performance

```{r}
augment(ridge_final_fit, new_data = Hitters_test) %>%
  rsq(truth = Salary, estimate = .pred)
```

And it performs fairly well given what we saw earlier.

## The Lasso

We will use the `glmnet` package to perform lasso regression. `parsnip` does not have a dedicated function to create a ridge regression model specification. You need to use `linear_reg()` and set `mixture = 1` to specify a lasso model. The `mixture` argument specifies the amount of different types of regularization, `mixture = 0` specifies only ridge regularization and `mixture = 1` specifies only lasso regularization. Setting `mixture` to a value between 0 and 1 lets us use both.

The following procedure will be very similar to what we saw in the ridge regression section. The preprocessing needed is the same, but let us write it out one more time. 

```{r}
lasso_recipe <- 
  recipe(formula = Salary ~ ., data = Hitters_train) %>% 
  step_novel(all_nominal_predictors()) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_predictors())
```

Next, we finish the lasso regression `workflow`.

```{r}
lasso_spec <- 
  linear_reg(penalty = tune(), mixture = 1) %>% 
  set_mode("regression") %>% 
  set_engine("glmnet") 

lasso_workflow <- workflow() %>% 
  add_recipe(lasso_recipe) %>% 
  add_model(lasso_spec)
```

While we are doing a different kind of regularization we still use the same `penalty` argument. I have picked a different range for the values of penalty since I know it will be a good range. You would in practice have to cast a wide net at first and then narrow on the range of interest.

```{r}
penalty_grid <- grid_regular(penalty(range = c(-2, 2)), levels = 50)
```

And we can use `tune_grid()` again.

```{r}
#| fig-alt: |
#|   Facetted connected scatter chart. regularization along the
#|   x-axis. Performance values along the y-axis. The facets are
#|   rmse and rsq. Both are fairly constant for low values of
#|   regularization, rmse starts moderately increasing and rsq
#|   starts moderately decreasing once the regularization
#|   gets larger.
tune_res <- tune_grid(
  lasso_workflow,
  resamples = Hitters_fold, 
  grid = penalty_grid
)

autoplot(tune_res)
```

We select the best value of `penalty` using `select_best()`

```{r}
best_penalty <- select_best(tune_res, metric = "rsq")
```

And refit the using the whole training data set.

```{r}
lasso_final <- finalize_workflow(lasso_workflow, best_penalty)

lasso_final_fit <- fit(lasso_final, data = Hitters_train)
```

And we are done, by calculating the `rsq` value for the lasso model can we see that for this data ridge regression outperform lasso regression.

```{r}
augment(lasso_final_fit, new_data = Hitters_test) %>%
  rsq(truth = Salary, estimate = .pred)
```




